---
title: "p8105_hw01_lg3158"
author: "Ling"
date: "9/23/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Problem 1
Create a dataframe comprised of:
  - a random sample of size 10 from a standard Normal distribution
  - a logical vector indicating whether elements of the sample are greater than 0
  - a character vector of length 10
  - a factor vector of length 10, with 3 different factor “levels”
```{r Problem 1-create df}
library(tidyverse)

p1 <- tibble(
  v_random = rnorm(10),
  v_pos = v_random > 0,
  v_char = as.character(c("a","b","c","d","e","f","g","h","i","j")),
  v_fac = as.factor(c("i","ii","iii","iii","ii","i","ii","iii","iii","iii"))
  )

knitr::kable(p1)

mean(pull(p1,v_random))
mean(pull(p1,v_pos))
mean(pull(p1,v_char))
mean(pull(p1,v_fac))

```
From the output above, only the mean of the random sample of size from a standard Normal distribution and the logic variable work. The other 2 don't work.

Transform the data type of the 3 variables:
```{r Problem 1-data type transformation, eval=FALSE}
as.numeric(pull(p1,v_char))
as.numeric(pull(p1,v_fac))
```
  From the console, the character variable cannot be transformed into numeric variables, but the factor vector can be transformed into 1, 2, 3.

Because the character variable is constructed by characters, but the factor vector has 3 levels ("i","ii","iii") that can be transformed into numbers. It partially helps explain the errors when calculating the mean because the factor vector with 3 levels cannot be transformed into numeric automatically before calculating the mean.

# Problem 2

Load the `penguins` dataset:

```{r Problem 2-load the dataset, eval=F, echo=T}
data("penguins", package = "palmerpenguins")
```
  - The data in this dataset include the variables as follow: `r data("penguins", package = "palmerpenguins")` `r names(penguins)`. The brief summary of the dataset is below. 
    `r skimr::skim(penguins)`
  
  - The dataset is a `r nrow(penguins)` x `r ncol(penguins)` dataframe.
  
  - The mean flipper length is `r mean(pull(penguins,flipper_length_mm),na.rm=T)` mm.

Make a scatterplot of `flipper_length_mm(y) vs. bill_length_mm(x)` in different `species`.
```{r Problem 2-scatterplot}
library(palmerpenguins) #add the package to make R markdown knit into Html
ggplot(data = penguins, mapping = aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point()

ggsave("p8105_hw1_lg3158_Q2scatter.tiff",plot = last_plot(), path = NULL)

```


